Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-11-21 Moltres.tex
Strong ans weak scalability plots and text in section 4.3 added

The present work introduces the open source simulation tool, Moltres
\cite{lindsay_moltres_2017}, for simulating \glspl{MSR}.  By implementing
deterministic neutronics and thermal hydraulics in the context of the
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group
neutron diffusion, temperature, and precursor governing equations in anywhere
from one to three dimensions and can be deployed on an arbitrary number of
processing units.

@@

\subsection{Scaling study}

As described earlier, Moltres runs naturally in parallel. To test its
performance, we conducted a scaling study with a typical 2D axisymmetric input
file. \Cref{fig:scaling_study} shows Moltres' performance. The solution
time follows the ideal closely; slight deviations are expected due to
the expense of communication and the nature of iterative solvers. More
parallelism means the solution process uses more ``older'' information, 
which leads to slower convergence in terms of iterations 
\cite{satish_balay_petsc_2015}.

\begin{figure}[htpb]
  \centering
  \includegraphics{scaling_study.eps}
  \caption{Moltres performance on a typical 2D axisymmetric problem for
    different numbers of processors}
  \label{fig:scaling_study}
\end{figure}

>>>>

The present work introduces the open source simulation tool, Moltres
\cite{lindsay_moltres_2017}, for simulating \glspl{MSR}.  By implementing
deterministic neutronics and thermal hydraulics in the context of the
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group
neutron diffusion, temperature, and precursor governing equations in anywhere
from one to three dimensions and can be deployed on an arbitrary number of
processing units. Through its computing power, Moltres is devoted to previously
unmatched fidelity in coupled neutronics and thermal hydraulics \gls{MSR}
simulation.

@@

\subsection{Scaling performance}

Parallelization in Moltres implemented via LibMesh, which includes a set of utilities
 for massively parallel finite element based computations, including mesh input/output,
 finite element library and interface for connections with solver packages. Employing 
the interfaces in LibMesh provides Moltres a significant flexibility including the ability 
to swap out solver libraries such as PetSc, which includes an expanding suite of parallel 
linear and nonlinear solvers. Problem domain decomposition utilized using LibMesh mesh 
adaptation capabilities for running on a specific number of processors and can either be 
performed manually before the start of the simulation or automatically at the parameters of computation.

In case of strong scaling, the problem size keeps fixed but the number of processors is increased. 
The goal of strong scaling study is to find an optimal ratio between a number of processors and elements 
that allows the computation to finish in a reasonable amount of time, and does not waste too much computational
 power due to parallel overhead. We measured Moltres performance for the strong scaling of a simple 2D 
axisymmetric case for various problem sizes separately for intra-node (2,820; 5,640; 11,280 and 28,200 elements) 
and extra-node (86,655; 173,310; 317,735; 664,355 elements) setup on Blue Waters' XK7 nodes (1 AMD 6276 Interlagos 
CPU, 16 integer scheduling units, 32 virtual cores, nominal clock speed 2.3 GHz).

\Cref{fig:intra_strong_scaling} shows Moltres performance, i.e. seconds per element vs. number of cores on 1 node
 (maximum 32 cores). As can be seen, up to 8 cores larger problem required considerably more time per element 
 because of cache overhead, but beyond this scaling demonstrates asymptotic dependence of the number of processors 
 due to increasing communication costs. The best parallel efficiency for intra-node study is approximately 89\% and 
 has been achieved for the largest problem (28,200 elements).

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{intra-node_strong.png}
  \caption{Moltres intra-node strong scaling efficiency for various problem size, for 1 \ldots 32 cores.}
  \label{fig:intra_strong_scaling}
\end{figure}

\Cref{fig:extra_strong_scaling} shows Moltres strong scaling up to 768 processors which takes into account communication costs 
between nodes. Similarly to intra-node study, up to 128 cores cache overhead causes performance slow down for a larger problem, 
while beyond 256 cores time per element stays almost constant for small cases (86,655 and 173,310 elements) and slighly 
decreasing for two larger problems. For extra-node scaling parallel efficiency also growth with the problem size and reach 
optimal value about 73\% for 664,355 elements.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=\maxwidth{\textwidth}]{extra-node_strong.png}
  \caption{Moltres extra-node strong scaling efficiency for various problem size, for 1 \ldots 24 nodes.}
  \label{fig:extra_strong_scaling}
\end{figure}

For the weak scaling study, the part of the problem (workload) assigned to each processor stays constant and additional elements are 
used to solve a bigger problem which would not fit in memory on a single node. Thus, the weak scaling measurement is justification
 for memory-bound application such as multiphysics code. Linear weak scaling is achieved when the execution time stays constant 
 while the workload increasing in direct proportion to the number of cores. We performed Moltres weak-scaling tests on Blue Waters, 
 keeping the workload constant at 581, 985, 1970 and 3940 elements per core. \Cref{fig:intra_weak} shows Moltres weak scaling 
performance measured for 1 \ldots 32 core within one Blue Waters node and \Cref{fig:extra_weak} demonstrates performance for 
32 \ldots 128 core. As expected, the largest drop in performance occurs when the number of cores increases from one to $\approx$ 8, which 
corresponds to switching from no communication to a 2-D domain decomposition. The further reduction in performance of only about 50\% over 
a range of 32 cores is likely caused by increased communication latency appearing from collective MPI calls. For extra-node case, performance
 drops about three times most likely because of poor node selection by job scheduler and significantly increased latency and bandwidth costs.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.75\textwidth]{intra-node_weak.png}
  \caption{Weak scaling performance of Moltres on Blue Waters, in seconds per element vs. number of processors, for a constant number of particles
per processor, for 1 \ldots 32 cores.}
  \label{fig:intra_weak}
\end{figure}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.75\textwidth]{extra-node_weak.png}
  \caption{Weak scaling performance of Moltres on Blue Waters, in seconds per element vs. number of processors, for a constant number of particles
per processor, for 32 \ldots 128 cores.}
  \label{fig:extra_weak}
\end{figure}

Moltres scalability study results clearly indicate that parallelization using LibMesh automatical domain decomposition is not perfectly efficient and
 further parallel performance optimization needed. Thus, focus on communication patterns study necessarily to identify latency and bandwidth costs
 and to optimize communications. Moreover, Moltres is memory-bound code and very sensitive to host memory and memory bandwidth, consequently, we will
 consider a transition from CPUs computing to GPU-accelerated computing which much faster for this type of algorithms because operates on the fly with global 
 memory instead of store in a CPU's cache before computations. Another way to improve parallel performance is to force the solver use more ``older'' 
 information from previous iteration, which leads to slower convergence in terms of iterations and workload increasing \cite{satish_balay_petsc_2015}.
